A Linear Solution of the Linearity Challenge

### 2025-12-17 paper/related.tex
more info

@@

>>>>

\Cref{tab:comparison} summarizes a comparison among the formalizations of binary
session calculus that we are aware of, roughly sorted by publication date. We
have selected those formalizations that define a calculus of binary sessions
equipped with a reduction/evaluation semantics with proofs of subject reduction
and possibly of deadlock freedom.
%
The formalizations differ substantially in terms of reference calculus, used
proof assistant, process/expression/channel representation, handling of
linearity.
%
Most importantly, almost all of them predates the publlication of the linearity
challenge~\cite{CarboneEtAl24} hence were not necessarily targeted to solving
that particular problem. Nonetheless, they provide a broad spectrum of
techniques

\fullcite{GotoEtAl16} describe the oldest formalization of a session-based
calculus for which we were able to access the source code (the formalization of
\citet{Gay01} predates this work, but the source code is no longer publicly
available).

fullcite{RouvoetPoulsenKrebbersVisser20} also formalize a subset of \CP as case
study for their approach based on separation logic, which we have partially
taken inspiration from for the representation of processes
(\cref{sec:context-agda,sec:process-agda}). Their formalization adopts the same
intrinsically-typed representation of processes with nameless channels as our
own, but their formalization focuses on the implementation of an interpreter (no
small-step semantics is defined) and subject reduction is

@@

\begin{table}[t]
    \footnotesize
    \newcommand{\textlabel}[1]{\rotatebox{90}{#1}}
    \newcommand{\YES}{$+$}
    \newcommand{\NO}{$-$}
    \newcommand{\MAYBE}{$\pm$}
    \caption{\label{tab:comparison} A summary of different formalizations of binary session calculi. All sizes are in kb.}
    \begin{minipage}{.9\textwidth}
    \begin{tabular}{@{}r@{~}|@{~}c@{~}|@{~}c@{~~}c@{~~}c@{~~}c@{~~}c@{~~}c@{~~}c@{~}|@{~}r@{~~}r@{~}|@{~}r@{~~}r@{~}|@{~}r@{~}l@{}}
        \textbf{Formalization} &
        \textlabel{Proof assistant} &
        \textlabel{Calculus} &
        \textlabel{Intrinsically typed} &
        \textlabel{Channel linearity} &
        \textlabel{Channel names} &
        \textlabel{Continuations} &
        \textlabel{Shared channels} &
        \textlabel{Polymorphism} &
        \textlabel{Subject reduction} &
        \textlabel{Library} &
        \textlabel{Deadlock freedom} &
        \textlabel{Library} &
        \textlabel{\bf Total size} &
        \\
        \hline
        \fullcite{GotoEtAl16} &
        Coq &
        \SessionPI &
        \NO &
        splitting &
        loc. nameless &
        \NO &
        \YES &
        \YES &
        543 &
        \NO &
        \NO &
        \NO &
        \textbf{543} &
        \footnote{Excluding safety results.}
        \\
        \fullcite{Thiemann19} &
        Agda &
        \SessionLAM &
        \YES &
        splitting &
        co-de Bruijn &
        \MAYBE &
        \YES &
        \NO &
        177 &
        \NO &
        \NO &
        \NO &
        \textbf{177} &
        \footnote{Including support for recursive types, subtyping and the interpreter.}
        \\
        \fullcite{CastroFerreiraYoshida20} &
        Coq &
        \SessionPI &
        \NO &
        splitting &
        loc. nameless &
        \NO &
        \YES &
        \NO &
        204 &
        \NO &
        \NO &
        \NO &
        \textbf{204} &
        \footnote{Excluding the original syntax.}
        \\
        \fullcite{CicconePadovani20} &
        Agda &
        \LinearPI & 
        \YES &
        splitting &
        co-de Bruijn &
        \YES &
        \YES &
        \NO &
        77 &
        \NO &
        \NO &
        \NO &
        \textbf{77} &
        \footnote{Including support for recursive and dependent session types.}
        \\
        \fullcite{RouvoetPoulsenKrebbersVisser20} &
        Agda &
        \CP &
        \YES &
        splitting &
        co-de Bruijn &
        \YES &
        \NO &
        \NO &
        27 &
        55 &
        \NO &
        \NO &
        \textbf{82} &
        \footnote{Including type-preserving evaluator and library for proof-relevant separation algebra.}
        \\
        \fullcite{ZalakainDardha21} &
        Agda &
        \GradedPI &
        \NO &
        leftovers &
        de Bruijn &
        \YES &
        \YES &
        \NO &
        82 &
        8 &
        \NO &
        \NO &
        \textbf{90} &
        \footnote{Including library for algebra of types.}
        \\
        \fullcite{JacobsBalzerKrebbers22} &
        Coq &
        \SessionLAM &
        \NO &
        splitting &
        named &
        \MAYBE &
        \NO &
        \NO &
        68 &
        \NO &
        25 &
        171 &
        \textbf{264} &
        \footnote{Including deadlock freedom and the library for connectivity graphs.}
        \\
        \fullcite{SanoKavanaghPientka23} &
        Beluga &
        \SCP &
        \NO &
        predicates &
        HOAS &
        \MAYBE &
        \NO &
        \NO &
        35 &
        \NO &
        \NO &
        \NO &
        \textbf{35}
        \\
        \fullcite{ZackonEtAl25} &
        Beluga &
        \CP &
        \NO &
        tags &
        HOAS &
        \NO &
        \NO &
        \NO &
        56 &
        73 &
        \NO &
        \NO &
        \textbf{129} &
        \footnote{Excluding the correspondence between \CP and \SCP.}
        \\
        this work &
        Agda &
        \Calculus &
        \YES &
        splitting &
        co-de Bruijn &
        \YES &
        \YES &
        \YES &
        21 &
        \NO &
        15 &
        \NO &
        \textbf{36} &
        \footnote{Including deadlock freedom and excluding safety results.}
    \end{tabular}
    \end{minipage}
\end{table}

paper di Marco a ECOOP (multiparty)
